---
title: "A Report on Lake Ontario's Microbes"
author: "Mark Watson"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Prepare the R environment

```{r load-packages}
#load libraries/packages for file
library(tidyverse)
```

# Load in the Lake Ontario Data

```{r load-data}
# load in lake ontario microbial community data
sample_and_taxon <-
  read_csv("data/sample_and_taxon.csv")
#inspect data
glimpse(sample_and_taxon)
sample_data <-
  read_csv("data/sample_data.csv")
```

# Lake Ontario Microbial Abundance Versus Temperature

```{r cells_v_temp}
# plot
# temp on x
# cel abundance on Y
# colored by env_group
# make it look nice
ggplot(data = sample_data) +
  aes(x = temperature) +
  labs(x = "Temperature (C)") +
  aes(y = cells_per_ml) +
  labs(y = "Cell Abundance (Cells/mL)")+
  geom_point(aes(color = env_group)) +
  labs(title = "Does temperature affect microbial abundance?") +
  geom_smooth(method = lm, color = "black") +
  labs(color = "Environmental Group") 
```

The above plot shows that:

- Temperature and Cell Abundance are positively correlated.
- Deep Samples are the coldest and have the fewest cells.
- Shallow Samples are warmer and have more cells.

```{r data-summary, include = FALSE}
n_samples <- n_distinct(sample_and_taxon$sample_id)


#Calculate the minimum and max temps
#min
sample_and_taxon %>%
  summarise(round(min(temperature), digits = 1)) %>%
  pull
#max
sample_and_taxon %>%
  summarise(round(max(temperature), digits = 1)) %>%
  pull()
```

The total number of samples is `r n_samples`. For this set of samples, temperature ranged from `r minTemp` to `r maxTemp` Degrees Celsius.